---
title: "initial exploration"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(janitor)
library(readr)
library(tibble)
library(dplyr)
```


```{r import cc institution data and clean}
ccid <- read_csv(here::here("data/cc_institution_details.csv")) 

ccid_var <- ccid %>%
  #removing unneeded variables
  select(-(contains("percentile"))) %>% 
  select(-(starts_with("vsa_"))) %>% 
  select(-site, -similar, -nicknames, -counted_pct, -cohort_size, -hbcu, 
         -awards_per_state_value, -awards_per_natl_value, -exp_award_state_value, 
         -exp_award_natl_value, -grad_150_value, -state_sector_ct, 
         -carnegie_ct) %>% 
  #convert flagship to binary.  
  mutate(flagship_binary = case_when(flagship == "X" ~ 1,
                                     flagship == "NULL" ~ 0),
  #convert percentages to decimal values
         awards_per_value = awards_per_value/100,
         ft_pct = ft_pct / 100,
         pell_value = pell_value / 100,
         ft_fac_value = ft_fac_value / 100,
         retain_value = as.numeric(retain_value) / 100)
```

```{r save the output as a csv}
write_excel_csv(x = ccid_var,
                path = "ccid_var.csv")
```


## Cleanup notes: 

* percetiles should be converted to decimals? making a list of these.
* binaries from `X` and `NULL` to `1` and `0`

## The data includes the following variables. I'm re-grouping them by the ones I think may provide value for the niche area we're studing. 

* `unitid` - Department of Education Unit ID number
* `chronname` - Institution name
* `city` - Institution city
* `state` - Institution state
* `level` - Level of institution (4-year, 2-year)           [[We might think of only keeping 4 year?]]
* `control` - Control of institution (Public, Private not-for-profit, Private for-profit)
* `basic` - Carnegie Foundation for the Advancement of Teaching Basic Classification (2010 version)
* `flagship` - Denotes Public flagship institutions  
* `flagship_binary` - This is the `flagship` variable converted to `1` and `0`.  
* `long_x` - Institution longitude                          
* `lat_y` - Institution latitude    
* `student_count` - Total number of undergraduates in 2010  [[maybe a measure of GIGO?]]
* `awards_per_value` - Completions per 100 FTE undergraduate students (average 2011, 2012, and 2013)  **This is our graduation rate.**
* `exp_award_value` - Estimated educational spending (expenses related to instruction, research, public service, student services, academic support, institutional support, operations and maintenance) per academic award in 2013. Includes all certificates and degrees
* `ft_pct` - Percentage of undergraduates who attend full-time  [[needs to convert to decimal]]
* `fte_value` - total number of full-time equivalent undergraduates 
* `med_sat_value` - Median estimated SAT value for incoming students
* `aid_value` - The average amount of student aid going to undergraduate recipients
* `endow_value` - End-of-year endowment value per full-time equivalent student
* `pell_value` - percentage of undergraduates receiving a Pell Grant  [[needs to convert to decimal]]
* `retain_value` - share of freshman students retained for a second year  [[needs to convert to decimal]]
* `ft_fac_value` - Percentage of employees devoted to instruction, research or public service who are full-time and do not work for an associated medical school  [[needs to convert to decimal]]

## Removed:

* hbcu - Denotes Historically Black Universities
* awards_per_state_value - Completions per 100 FTE undergraduate students, state and sector average maybe unnecessary - compares colleges, not students.
* awards_per_natl_value - Completions per 100 FTE undergraduate students, national sector average maybe unnecessary - compares colleges, not students.
* exp_award_state_value - Spending per completion, state average    maybe unnecessary - compares colleges, not students.
* exp_award_natl_value - Spending per completion, national average  maybe unnecessary - compares colleges, not students.
* med_sat_percentile - Institution's percent rank for median SAT value within sector maybe unnecessary - compares colleges, not students.
* aid_percentile - Institution's percent rank for average amount of student aid going to undergraduate recipients within sector  maybe unnecessary
* endow_percentile - Institution's percent rank for endowment value per full-time equivalent student within sector maybe unnecessary
* grad_100_value - percentage of first-time, full-time, degree-seeking undergraduates who complete a degree or certificate program within 100 percent of expected time (bachelor's-seeking group at 4-year institutions)
* grad_100_percentile - Institution's percent rank for completers within 100 percent of normal time within sector
* grad_150_value - percentage of first-time, full-time, degree-seeking undergraduates who complete a degree or certificate program within 150 percent of expected time (bachelor's-seeking group at 4-year institutions)
* grad_150_percentile - Institution's percent rank for completers within 150 percent of normal time within sector
* pell_percentile - Institution's percent rank for undergraduate Pell recipients within sector maybe unnecessary
* retain_percentile - Institution's percent rank for freshman retention percentage within sector
* ft_fac_percentile - Institution's percent rank for full-time faculty share within sector maybe unnecessary
* vsa_year - Most recent year of Student Success and Progress Rate data available
* vsa_grad_after4_first - First-time, full-time students who graduated from this institution within four years
* vsa_grad_elsewhere_after4_first - First-time, full-time students who graduated from another institution within four years   --- Did they LEAVE to graduate? Maybe unnecessary, but good context.
* vsa_enroll_after4_first - First-time, full-time students who are still enrolled at this institution after four years maybe unnecessary
* vsa_enroll_elsewhere_after4_first - First-time, full-time students who are enrolled at another institution after four years
* vsa_grad_after6_first - First-time, full-time students who graduated from this institution within six years Maybe unnecessary, but good context.
* vsa_grad_elsewhere_after6_first - First-time, full-time students who graduated from another institution within six years Maybe unnecessary, but good context.
* vsa_enroll_after6_first - First-time, full-time students who are still enrolled at this institution after six years Maybe unnecessary, but good context.
* vsa_enroll_elsewhere_after6_first - First-time, full-time students who are enrolled at another institution after six years Maybe unnecessary, but good context.
* vsa_grad_after4_transfer - Full-time transfer students who graduated from this institution within four years maybe unecessary
* vsa_grad_elsewhere_after4_transfer - Full-time transfer students who graduated from another institution within four years maybe unecessary
* vsa_enroll_after4_transfer - Full-time transfer students who are still enrolled at this institution after four years maybe unecessary
* vsa_enroll_elsewhere_after4_transfer - Full-time transfer students who are enrolled at another institution after four years maybe unecessary
* vsa_grad_after6_transfer - Full-time transfer students who graduated from this institution within six years
* vsa_grad_elsewhere_after6_transfer - Full-time transfer students who graduated from another institution within six years
* vsa_enroll_after6_transfer - Full-time transfer students who are still enrolled at this institution after six years
* vsa_enroll_elsewhere_after6_transfer - Full-time transfer students who are enrolled at another institution after six years
* counted_pct - share entering undergraduate class who were first-time, full-time, degree-seeking students, meaning that they generally would be part of a tracked cohort of potential graduates. The entering class of 2007 is displayed for 4-year institutions; 2010 for 2-year institutions. 


```{r}
glimpse(ccid_var)

ccid_var %>% 
  filter(flagship_binary == 1) %>% 
  arrange(state)

glimpse(ccid_var %>% 
  filter(chronname == "Bellevue College"))

ccid_like_bc <- ccid_var %>%
  filter(level == "4-year",
         control == "Public",
         flagship_binary == 0) 

write_excel_csv(x = ccid_like_bc,
                path = "ccid_like_bc.csv")
```
